Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-07 tests-sodshocktube.tex
minor text edits

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/SodShockTube.eps}
\caption{The density distribution of the classic Sod Shock Tube for
three different solvers (from left to right column) and with (bottom
row) and without (top row) AMR.  In each case 100 zones are used on
the root level and the results are shown at $t=0.25$.  All cells are
plotted and color-coded by level (at the time shown, only small region
surrounding the contact discontinuity and the shock are refined).  In
each panel, we show the analytic solution as a solid line and the $E_1$
and $E_\infty$ error norms in the upper right.}
\label{fig.sodshocktube}
\end{center}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/SodShockTube.eps}
\caption{The density distribution of the classic Sod Shock Tube for
three different solvers (from left to right column) and with (bottom
row) and without (top row) AMR.  In each case 100 zones are used on
the root level and the results are shown at $t=0.25$.  All cells are
plotted and color-coded by level with purple 
indicating level 0, green level 1 and red level 2 (at the time shown, only small region
surrounding the contact discontinuity and the shock are refined).  In
each panel, we show the analytic solution as a solid line and the $E_1$
and $E_\infty$ error norms in the upper right.}
\label{fig.sodshocktube}
\end{center}
\end{figure}

### 2013-06-07 tests-vandv.tex
minor text edits

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code in a variety of
set-ups, including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
Compared to the other codes involved in these projects, the \enzo\ code typically has a more difficult time resolving
small-scale self-gravitating structures (for an equivalent dark matter
particle count and nominally equivalent force resolution), but does
comparably well as a tree-based code for larger structure, and is
typically superior in terms of resolving fluid features due to its
higher-order (and artificial viscosity-free) Piecewise Parabolic
Method hydrodynamics solver.  When examining classical test cases such
as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
clusters with very similar density, temperature, and entropy profiles
to other grid-based codes that use Godunov-type hydro methods, which
systematically differed from particle-based codes using SPH in this
code comparison.  Similarly, in galaxy cluster simulations that look
specifically at the properties of cosmological shocks
\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
similar to other high-order grid-based hydrodynamics codes, and a far
superior performance of \enzo\ is observed (in terms of resolution of
fluid features and shock detection) in low-density regions when
compared to a particle-based code.  In tests of the Lyman-alpha forest
that include radiative cooling and a uniform metagalactic ultraviolet
background, \enzo\ and Gadget provide results on metrics such as the
matter power spectrum that are comparable to within 5\%
\citep{2007MNRAS.374..196R}.

@@

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009},
which involved tests with and without analytical solutions.  \enzo\ produced
results similar to both the analytical solutions and results obtained by 
other codes.  We note, however, that there were minor differences throughout
the comparison between various codes, and the majority of the codes differed
in at least a subset of the tests.
\citet{2011ApJ...726...55T} shows the results of varied reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and Gadget-2 codes, observing similar trends between
the two codes. However, at nominally equivalent resolution
where the particle and cell gas masses are comparable, 
\enzo\ simulations typically displayed a
substantially higher level of gas structure.  This is unsurprising due
to \enzo's higher-order hydro solver.  Finally,
\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
the context of common-envelope binary stellar interactions.  The
authors show that the codes display reasonable convergence properties
as a function of simulation resolution, and also agree quite well with
each other. However, the observed mass-loss rates do not agree
particularly well with observations.

>>>>

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code in a variety of
set-ups, including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
Compared to the other codes involved in these projects, the \enzo\ code typically has a more difficult time resolving
small-scale self-gravitating structures (for an equivalent dark matter
particle count and nominally equivalent force resolution), but does
comparably well as a tree-based code for larger structure, and is
typically superior in terms of resolving fluid features due to its
higher-order (and artificial viscosity-free) Piecewise Parabolic
Method hydrodynamics solver.  When examining classical test cases such
as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
clusters with very similar density, temperature, and entropy profiles
to other grid-based codes that use Godunov-type hydro methods, which
systematically differed from particle-based codes using SPH in this
code comparison.  Similarly, in galaxy cluster simulations that look
specifically at the properties of cosmological shocks
\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
similar to other high-order grid-based hydrodynamics codes, and a far
superior performance of \enzo\ is observed (in terms of resolution of
fluid features and shock detection) in low-density regions when
compared to a particle-based code.  In tests of the Lyman-alpha forest
that include radiative cooling and a uniform metagalactic ultraviolet
background, \enzo\ and GADGET provide results on metrics such as the
matter power spectrum that are comparable to within 5\%
\citep{2007MNRAS.374..196R}.

@@

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009},
which involved tests with and without analytical solutions.  \enzo\ produced
results similar to both the analytical solutions and results obtained by 
other codes.  We note, however, that there were minor differences throughout
the comparison between various codes, and the majority of the codes differed
in at least a subset of the tests.
\citet{2011ApJ...726...55T} shows the results of varied reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and GADGET-2 codes, observing similar trends between
the two codes. However, at nominally equivalent resolution
where the particle and cell gas masses are comparable, 
\enzo\ simulations typically displayed a
substantially higher level of gas structure.  This is unsurprising due
to \enzo's higher-order hydro solver.  Finally,
\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
the context of common-envelope binary stellar interactions.  The
authors show that the codes display reasonable convergence properties
as a function of simulation resolution, and also agree quite well with
each other. However, the observed mass-loss rates do not agree
particularly well with observations.

### 2013-06-07 tests-wavepool.tex
minor text edits

In this one-dimensional test we pass a short wavelength linear wave through a static refined region.  The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.  We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.  The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.  The (unrefined) root grid is covered with 100 cells, resulting in a wavelength for the linear wave of only 10 cells.  This is, therefore, a challenging problem for hydro methods.  We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.

@@

Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations. However,
because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.

>>>>

In this one-dimensional test we pass a short wavelength linear wave through a static singley-refined region.  The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.  We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.  The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.  The (unrefined) root grid is covered with 100 cells, resulting in a wavelength for the linear wave of only 10 cells.  This is, therefore, a challenging problem for hydro methods.  We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.

@@

Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although with mild damping on exit. However, because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.
